Python as a machine learning language: experiences from the scikit-learn project

### 2013-06-11 paper.tex
elaborate on effect of array implementation to API
Also, at least mention the online learning algorithms that we have.

In \sklearn, we chose a representation of data that is as close as
possible to the matrix representation: datasets are encoded as NumPy
multidimensional arrays for dense data and as SciPy sparse matrices for sparse
data. While those may seem rather unsophisticated data representations when
compared to more object-oriented constructs, such as those used by
Weka \citep{hall2009weka}, they bring the prime advantage of allowing us to rely
% Lars: "orders of magnitude faster" is an implementation detail
on all NumPy and SciPy vector operations -- often orders of magnitude faster
that the corresponding Python loops -- while allowing us at the same time to
keep the code short and readable.  This design choice has also been motivated by
the fact that, given their pervasiveness in many other scientific Python
packages, many scientific users of Python are already familiar with NumPy dense
arrays and SciPy sparse matrices, hence helping them to familiarize with the
project.  From a practical point of view, these formats also provide a bunch of
data loading and conversion tools which make them very easy to use in many
contexts. Moreover, for tasks where the inputs are text files or semi-structured
objects, we provide \textit{vectorizer} objects that efficiently convert such
data to the NumPy or SciPy formats.

>>>>

In \sklearn, we chose a representation of data that is as close as
possible to the matrix representation: datasets are encoded as NumPy
multidimensional arrays for dense data and as SciPy sparse matrices for sparse
data. While those may seem rather unsophisticated data representations when
compared to more object-oriented constructs, such as those used by
Weka \citep{hall2009weka}, they bring the prime advantage of allowing us to rely
% Lars: "orders of magnitude faster" is an implementation detail
on all NumPy and SciPy vector operations while allowing us at the same time to
keep the code short and readable.  This design choice has also been motivated by
the fact that, given their pervasiveness in many other scientific Python
packages, many scientific users of Python are already familiar with NumPy dense
arrays and SciPy sparse matrices, hence helping them to familiarize with the
project.  From a practical point of view, these formats also provide a bunch of
data loading and conversion tools which make them very easy to use in many
contexts. Moreover, for tasks where the inputs are text files or semi-structured
objects, we provide \textit{vectorizer} objects that efficiently convert such
data to the NumPy or SciPy formats.

The public API is oriented toward batch processing
to enable efficient implementation.
While classification and regression algorithms
can make predictions for single samples,
and some algorithms in \sklearn can perform online learning
(e.g. stochastic gradient descent for linear SVMs and logistic regression,
an online variant of $k$-means based on \citealt{sculley2010web}),
these algorithm are really intended to be used on small batches of samples.
Doing so makes optimal use of NumPy and SciPy
because they prevent the overhead inherent in Python function calls
and the language's per-element dynamic type checking.
